Streaking and Clutching in Round-Based First Person Shooter (FPS) Esports

### 2022-07-16 presentation.qmd
add b2b win props

```{r}
#| label: cod_heatmap_prep
summarize_rounds <- function(rounds, ...) {
  rounds |> 
    group_by(pre_cumu_w, pre_cumu_l, ...) |> 
    summarize(
      n = n(),
      across(c(win_round, win_series), sum)
    ) |> 
    ungroup() |> 
    mutate(
      win_round_prop = win_round / n,
      win_series_prop = win_series / n
    )
}

n_rounds_side <- rounds |> summarize_rounds(is_offense)
o_win_prop <- n_rounds_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

@@

common_heatmap_layers <- function(...) {
  list(
    ...,
    guides(
      fill = 'none'
    ),
    scale_x_continuous(
      labels = 0:5,
      breaks = seq(.5, 5.5, by = 1),
      expand = c(0, 0)
    ),
    scale_y_continuous(
      labels = 0:5,
      breaks = seq(.5, 5.5, by = 1),
      expand = c(0, 0),
      sec.axis = sec_axis(
        trans = I, 
        name = ' ', 
        breaks = seq(0.5, 5.5, by = 1), 
        labels = rep('', 6)
      )
    ),
    theme(
      panel.grid.major = element_blank(),
      axis.title = element_text(hjust = 0.5),
      plot.title = ggtext::element_markdown(hjust = 0.5),
      axis.text = element_text(size = 16, face = 'bold')
    ),
    labs(
      caption = 'CDL SnD Major and Weekly matches, 2020 - present',
      tag = '**Viz**: @TonyElHabr | **Data**: @IOUTurtle',
      x = "Offensive Team's # of Pre-Round Wins",
      y = "Defensive Team's # of Pre-Round Wins"
    )
  )
}

.dpi <- 300
save_heatmap <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 6, 
    height = 6,
    units = 'px',
    dpi = .dpi,
    ...
) {
  ggsave(
    plot = p,
    filename = filename,
    width = width * dpi,
    height = height * dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

heatmap_tile <- function(..., .xmin, .xmax, .ymin, .ymax, .fill) {
  list(
    ...,
    geom_rect(
      color = blackish_background,
      aes(
        fill = .data[[.fill]],
        xmin = .data[[.xmin]], 
        ymin = .data[[.ymin]],
        xmax = .data[[.xmin]], + 1,
        ymax = .data[[.ymin]] + 1
      )
    )
  )
}

pre_cumu_heatmap_tile <- function(..., .fill) {
  heatmap_tile(
    .xmin = 'pre_cumu_w',
    .ymin = 'pre_cumu_l',
    .fill = .fill,
    ...
  )
}

@@

add_text_layers <- function(..., .f) {
  list(
    ...,
    .f(.emphasize = 'low'),
    .f(.emphasize = 'high')
  )
}
```

```{r}
#| label: p_o_round_win_prop
add_o_round_win_text_layer <- function(..., .emphasize = 'low') {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'win_round_prop') +
  add_text_layers(.f = add_o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_heatmap(p_o_round_win_prop, width = 7.5, height = 5)
```

@@

add_o_round_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- 'white'
  } else {
    op <- `>=`
    .color <- blackish_background
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_round_prop',
      .color = .color,
      .threshold = -0.1,
      .labeller = heatmap_div_labeller
    )
  )
}

@@

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_round_prop') +
  add_text_layers(.f = add_o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_heatmap(p_o_round_win_prop_over_expected, width = 7.5, height = 5)
```

@@

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'win_series_prop') +
  add_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_heatmap(p_o_series_win_prop, width = 7.5, height = 5)
```

@@

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_series_prop') +
  add_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_heatmap(p_o_series_win_prop_over_expected, width = 7.5, height = 5)
```

@@

::: notes
-   There should be evidence of 6-0's and 6-5's being more frequent than expected given 50-50% chances. Might need to exclude major matchups, so that series are basically random.
:::

@@

pre_cumu_heatmap_tile <- function(..., .fill) {
  heatmap_tile(
    .xmin = 'n_rounds',
    .ymin = 'round',
    .fill = .fill,
    ...
  )
}

## now accounting for number of rounds that end up being played...
## seems like there is interesting behavior in the last round, regardless `n_rounds`
b2b_round_win_props <- rounds |> 
  filter(n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds)

b2b_round_win_props |> 
  ggplot() +
  pre_cumu_heatmap_tile(.fill = 'prop')

b2b_round_win_props |> 
  filter(n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds) |> 
  group_by(n_rounds) |> 
  mutate(
    intra_prop = (n - mean(n)) / sd(n)
  ) |> 
  ggplot() +
  aes(x = n_rounds, y = round) +
  geom_tile(aes(fill = intra_prop))

## now accounting for number of rounds that end up being played AND series winners...
rounds |> 
  filter(win_series, n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds)

## and the losers...
rounds |> 
  filter(!win_series) |> 
  summarize_b2b_round_wins(n_rounds)
```

>>>>

```{r}
#| label: cod_heatmap_prep
n_rounds_side <- rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

o_win_prop <- n_rounds_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

@@

common_cod_labs <- function(...) {
  list(
    ...,
    labs(
      caption = 'CDL SnD Major and Weekly matches, 2020 - present',
      tag = '**Viz**: @TonyElHabr | **Data**: @IOUTurtle'
    )
  )
}

common_cod_heatmap_layers <- function(...) {
  list(
    ...,
    guides(
      fill = 'none'
    ),
    scale_x_continuous(
      labels = 0:5,
      breaks = seq(0.5, 5.5, by = 1),
      expand = c(0, 0)
    ),
    scale_y_continuous(
      labels = 0:5,
      breaks = seq(0.5, 5.5, by = 1),
      expand = c(0, 0),
      sec.axis = sec_axis(
        trans = I, 
        name = ' ', 
        breaks = seq(0.5, 5.5, by = 1), 
        labels = rep('', 6)
      )
    ),
    theme(
      panel.grid.major = element_blank(),
      axis.title = element_text(hjust = 0.5),
      plot.title = ggtext::element_markdown(hjust = 0.5),
      axis.text = element_text(size = 16, face = 'bold')
    ),
    common_cod_labs(),
    labs(
      x = "Offensive Team's # of Pre-Round Wins",
      y = "Defensive Team's # of Pre-Round Wins"
    )
  )
}

.dpi <- 300
save_heatmap <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 7.5, 
    height = 5,
    units = 'px',
    dpi = .dpi,
    ...
) {
  ggsave(
    plot = p,
    filename = filename,
    width = width * dpi,
    height = height * dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

heatmap_tile <- function(..., .fill) {
  list(
    ...,
    geom_rect(
      color = blackish_background,
      aes(
        fill = .data[[.fill]],
        xmin = pre_cumu_w, 
        ymin = pre_cumu_l,
        xmax = pre_cumu_w + 1,
        ymax = pre_cumu_l + 1
      )
    )
  )
}

@@

heatmap_text_layers <- function(..., .f) {
  list(
    ...,
    .f(.emphasize = 'low'),
    .f(.emphasize = 'high')
  )
}
```

```{r}
#| label: p_o_round_win_prop
o_round_win_text_layer <- function(..., .emphasize = 'low') {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(.f = o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_heatmap(p_o_round_win_prop)
```

@@

o_round_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- 'white'
  } else {
    op <- `>=`
    .color <- blackish_background
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_round_prop',
      .color = .color,
      .threshold = -0.1,
      .labeller = heatmap_div_labeller
    )
  )
}

@@

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(.f = o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_heatmap(p_o_round_win_prop_over_expected)
```

@@

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_heatmap(p_o_series_win_prop)
```

@@

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_heatmap(p_o_series_win_prop_over_expected)
```

@@

::: notes
-   There should be evidence of 6-0's and 6-5's being more frequent than expected given 50-50% chances.
:::

@@

## now accounting for number of rounds that end up being played...
## seems like there is interesting behavior in the last round, regardless `n_rounds`
b2b_round_win_props <- rounds |> 
  filter(n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds) |> 
  group_by(n_rounds) |> 
  mutate(
    z = (n - mean(n)) / sd(n),
    rnk = row_number(desc(n))
  ) |> 
  ungroup()

b2b_round_win_props_tile_layer <- function(..., .round, .palette, .is_new = TRUE) {

  init <- if(!.is_new) {
    list()
  } else {
    ggnewscale::new_scale_fill()
  }

  list(
    ...,
    init,
    geom_tile(
      data = b2b_round_win_props |> filter(n_rounds == .round),
      color = blackish_background,
      height = 0.8,
      show.legend = FALSE,
      aes(
        fill = z,
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    ),
    ggsci::scale_fill_material(palette = .palette) 
  )
}

b2b_round_win_prop_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `>`
    .color <- blackish_background
  } else {
    op <- `==`
    .color <- 'white'
  }
  list(
    ...,
    geom_text(
      data = b2b_round_win_props |> filter(`op`(rnk, 1)),
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        label = sprintf('%s\n(%s)', scales::percent(prop, accuracy = 1), scales::comma(n)),
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    )
  )
}

b2b_round_win_prop_text_layers <- function(...) {
  list(
    ...,
    b2b_round_win_prop_text_layer(.emphasize = 'low'),
    b2b_round_win_prop_text_layer(.emphasize = 'high')
  )
}

b2b_round_win_props_arws <- tibble(
  x = c(9.3, 9.4, 9.75, 10),
  y = c(8.25, 8.5, 8.75, 8.75),
  xend = c(8.1, 9.1, 9.5, 10.5),
  yend = c(7.5, 8.5, 9, 10),
  curvature = c(-0.2, -0.1, 0.1, 0.2)
) |> 
  mutate(
    rn = row_number()
  )

p_b2b_round_win_props <- b2b_round_win_props |> 
  ggplot() +
  b2b_round_win_props_tile_layer(.round = 7, .palette = 'pink', .is_new = FALSE) +
  b2b_round_win_props_tile_layer(.round = 8, .palette = 'light-blue') +
  b2b_round_win_props_tile_layer(.round = 9, .palette = 'light-green') +
  b2b_round_win_props_tile_layer(.round = 10, .palette = 'amber') +
  b2b_round_win_props_tile_layer(.round = 11, .palette = 'deep-orange') +
  b2b_round_win_prop_text_layers() +
  scale_x_continuous(
    labels = 2:11,
    breaks = seq(2.5, 11.5, by = 1),
    expand = c(0, 0.5)
  ) +
  scale_y_continuous(
    labels = 7:11,
    breaks = seq(7.5, 11.5, by = 1),
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  common_cod_labs() +
  labs(
    title = 'Back-to-back round win %',
    x = 'Round',
    y = '# of Rounds in Complete Series'
  ) + 
  geom_text(
    data = tibble(
      label = 'Back-to-back wins occur most\nfrequently in the last round.',
      x = 9.5,
      y = 8.5
    ),
    size = 9 / .pt,
    fontface = 'bold',
    color = 'white',
    hjust = 0,
    vjust = 0.5,
    aes(x = x, y = y, label = label)
  ) +
  map(
    b2b_round_win_props_arws |> group_split(rn), 
    function(.x) {
      geom_curve(
        data = data.frame(),
        color = 'white',
        arrow = arw,
        aes(x = .x$x, y = .x$y, xend = .x$xend, yend = .x$yend),
        curvature = .x$curvature
      )
    }
  )
save_heatmap(p_b2b_round_win_props)
# ## now accounting for number of rounds that end up being played AND series winners...
# rounds |> 
#   filter(win_series, n_rounds > 6L) |> 
#   summarize_b2b_round_wins(n_rounds)
# 
# ## and the losers...
# rounds |> 
#   filter(!win_series) |> 
#   summarize_b2b_round_wins(n_rounds)
```

![](b2b_round_win_props.png){fig-align="center"}
